Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-11-14 writeups/psycMethods/psycMethods.tex
foo

\thispagestyle{empty}
\section{Acknowledgements}

I would first like to thank my advisor, Mike Byrne, for all of his feedback and suggestions for this dissertation research
and also his excellent guidance throughout my time as a graduate student.
I would also like to thank my parents, Mike and Patsy, for the positive influence that they have provided me throughout my life.
Finally, I would like to thank my brothers Brendan and Justin, sister Kellen, extended family, and friends, for all of their support during my time in the Rice Psychology graduate program.

This work was supported in part by the Gertrude Maurin Fund.
The resources were used to purchase high-performance computing hardware and the research would not have been possible without this crucial equipment.

I would also like to thank the creators of a few open-sourced software packages that were heavily leveraged to do this research:
The majority of the modeling and analysis was done with the R statistical computing environment \parencite{RCoreTeam2014} and two important R packages:
data.table \parencite{Dowle2014} was leveraged to implement the models efficiently, and ggplot2 \parencite{Wickham2009} was used to produce the visualizations.

@@

\thispagestyle{empty}
\tableofcontents
\newpage

\thispagestyle{empty}
\listoftables
\newpage

\thispagestyle{empty}
\listoffigures
\newpage

@@

The core research question motivating this proposal is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
This process can be framed as a memory retrieval problem.
The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
That is, each user on a social media site has created content that contains a set of tags within the context of each post.

@@

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample1-crop.pdf}}}
    \hfill
    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample3-crop.pdf}}}
    \hfill
    \vfill
    \hfill
    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample4-crop.pdf}}}
    \hfill
    \subfigure[]{\fbox{\includegraphics[width=8cm]{twitterTweetExample5-crop.pdf}}}
    \hfill
    \caption{Example tweets on twitter.com}
    \label{figTweetExample}
  }%
\end{figure}

Some example tweets on Twitter are included in Figure \ref{figTweetExample}.
Note how users both add hashtags to the end of a tweet as a summary of the message and emphasize important concepts words in the tweet by turning them into hashtags.
That is, hashtags can be both intermixed within the words of the tweet and added to the end of the message.

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{stackoverflowPostExample1-crop.pdf}}}
  \caption{Example post on stackoverflow.com}
  \label{figSOExample}
\end{figure}

An example StackOverflow post is included in Figure \ref{figSOExample}.
The site requires that for each post the user must associate at least one tag with the post, and the average number of tags per post is around 3.
The author used 4 tags for this post: \emph{javascript}, \emph{html}, \emph{internet-explorer}, and \emph{xhtml}.

@@

\subsubsection{Lifetime of a Hashtag for a Specific User}

Prior research has examined the growth and decay cycle of hashtag use across users \parencite{Tsur2012}.
However, much less is known about hashtag lifetime for individual users.
It may not necessarily follow that a hashtag's life cycle for individual users matches the life cycle across users.
Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.

\subsubsection{User-Customized Hashtag Prediction}

The end goal of this work is to identify a memory retrieval model that can suggest relevant hashtags to users when they wish to retrieve one.
These hashtags should be customized to their specific interests and relevant to the content of the post that they are currently creating.
This will undoubtedly require a combination of two primary model components:
[1] a user's prior likelihood of choosing a particular hashtag, given their previous hashtag use, and [2] the likelihood that a particular hashtag is related to the context of the post being created.

A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is effectively infinite.
In these domains, the model can utilize a user's hashtag history as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
The end goal is to have a model that can properly balance the user's tag history with the contextual cues in the post in order to generate an accurate prediction.

@@

Using this assumption for large datasets ($S_{ji} = log \frac{p(i|j)}{p(i)}$), the interpretation of the context-adjusted odds ratio becomes much simpler.
If this ratio for a particular tag and context chunk is greater than one, then the log is positive, which means that this context has been observed more often with this tag than in general.
If this ratio is less than one, then the log is negative, which provides a negative adjustment of total activation since this context has been observed more often in general than with this particular tag.

@@

\subsubsection{Scaling the Equations}

The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
As an applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.

However, scale is also a theoretical concern.
The declarative memory system is tasked to quickly return a single chunk of information that is stored alongside millions to billions of other unique chunks given only the current context and past history of use.
To do this task, it seems plausible that a model of declarative memory must use a representation that produces accurate retrievals even when large amounts of observations are included.
Further, it seems plausible that the time to compute an activation value by a computer is correlated with the time to compute an activation value by the brain.
So any model of declarative memory should also be reasonably computationally efficient even with large co-occurrence datasets.

ACT-R's strength of association ($S_{ji}$) and the equivalent pointwise mutual information index ($\mathit{PMI}_{ji}$) have been shown to scale well for large corpora.
\textcite{Douglass2010} implemented the large-scale version of $S_{ji}$ in Erlang.
The wall-clock time required for a declarative memory retrieval request scaled linearly with the size of declarative memory with this implementation.
Retrieval times were around one second with the largest declarative memory that contained over one million associations (co-occurrence pairs).

The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} was developed to predict user navigation of web-based hyperlinks (e.g., choosing between a set of results from a search query).
The theory uses the large-scale version of $S_{ji}$ to measure the relatedness between a user's current goals and the current context.
It was shown to scale well when working with a large-scale search query database for internet search.

As one of the largest tests for the PMI index, \textcite{Farahat2004} generated the co-occurrence matrix from a 118GB local database of 10 million pages of the Stanford WebBase Project.
\citeauthor{Farahat2004} compared strength of association values using the PMI index and similarity ratings derived from human raters.
Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.

@@

\subsubsection{Retrieval Process}

Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
Decoding takes the memory vector for context and decodes it back into an environment vector.
The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.

Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.

The resonance and decoding processes are inversions of each other.
Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.

@@

For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
Within this corpus, the phrase ``martin luther'' often occurred.
When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated by convolving a 1-back location vector with luther ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin with and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.

Using decoding, the word-order memory vector for martin is deconvolved to the right ($o_{martin} \circledast \Phi$) which returns the environment vector for luther ($e_{luther}$).
Using resonance, a memory vector for ``martin $\Phi$'' is constructed ($e_{martin} * \Phi$), which correlates highest with the word-order memory vector for luther ($o_{luther}$).
Either way, luther is returned as the most likely term to match the phrase ``martin $\Phi$.''

New vectors generated by circular convolution are uncorrelated with all other environment vectors and have the same length.
This allows the word order information and unordered context information for a chunk to be aggregated and represented in the same memory vector.
Quite simply, the different types of memory vectors can be summed to create a single memory vector representation for each chunk (i.e., $m_{i} = c_{i} + o_{i}$).

@@

This is a counterintuitive result.
It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
This does not seem to be the case in testing however \parencites{Sahlgren2008,Jones2007}.
Nonetheless, \textcite{Kanerva2000} suggested to perform an SVD reduction technique on a compressed random permutation vector.
This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.

@@

Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.

The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
\textcite{Rutledge2008} used \textcite{Anderson1974}'s classic experimental paradigm to test the fan effect.
In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
They are later asked to recall if a particular person and place pair was presented to them before.
The main manipulation is the ``fan'' of the person and place for each pair.
The fan is defined as the number of different contexts that each item in the pair is presented in.
For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store,'' then ``hippie'' would have a fan of two.
Recall accuracy decrease and time to recall increase as the sum total fan of the person and place increases.

Both the ACT-R declarative memory theory and vector-based models produce the fan effect due to the effect that fan has on the co-occurrence matrix.
As a particular word (e.g., ``hippie'') is seen in more and more contexts, the co-occurrence count of the word across all contexts is spread across multiple contexts.
This means that when a retrieval for a context given the word occurs, multiple contexts have relatively high activation.
This makes it more difficult to test if a signal (e.g., if a word appears in a specific context) is higher than the background noise (e.g., all other contexts that the word has appeared in),
which makes the retrieval more uncertain.

Both vector-based and Bayesian models produce the fan effect since they both are based on co-occurrences.
However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.

@@

\subsection{Recommendation Systems}

A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
General recommendation systems have grown in need with the growth of the web.
Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.

\subsubsection{The Netflix Prize}

The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
Netflix is an online subscription service where users can stream television episodes and movies.
Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
In order to provide a high quality user experience on the site, Netflix created a content recommendation service.
This tool recommends to the user movies and television episodes that most likely match their specific interests.

Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension. 
This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
By using a factor analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.

\subsubsection{Recommending Followers on Twitter}

Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
The profiles were created by generating a word frequency matrix of words by (user + followers + followees), and then providing that matrix as input into the Lucene text search engine framework.

Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
It uses the term frequency-inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
The term frequency component measures the strength of association between a word and a document.
The inverse document frequency component measures how much information is contained within a specific word, independent of document.

When \citeauthor{Hannon2010}'s system recommended a set of individuals for a user to follow, it collected all of the words used in posts from that user, constructed a search query, and provided that query to Lucene.
Lucene then returned the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.

\subsubsection{Hashtags for Recommendation Systems}

\textcite{Efron2010} showed how hashtags might be used as a type of recommendation system for Twitter.
Users constructed a query of text that represented a search for a topic that they were interested in.
The system then retrieved the most likely hashtags that were associated with that search query.
The model is based on building a word co-occurrence matrix of words by hashtags, and then uses that matrix to assign a similarity score for each hashtag when provided with a search query.
Participants graded how relevant the hashtags returned were to each search query.
The results suggest that hashtags contain useful information that help classify the gist of a post. 
Although this model does not predict the specific hashtags used when composing a tweet, a model like this could certainly be tailored for that purpose.

@@

Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
Recommendation models for StackOverflow, Google+, and Twitter will be described in more detail to follow.

@@

It is interesting that in both studies the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
(i.e., when predicting the user-chosen tags for human-created content on social media sites).

\subsubsection{StackExchange}

A tag suggestion tool has been deployed for the \url{tex.stackexchange.org} \parencite{LatexTags2013} and \url{superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
Latex, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.

\begin{figure}[!htbp]
  \scalebox{.9}{\resizebox{\linewidth}{!}{\includegraphics{superUserTagSuggestion-crop.pdf}}}
  \caption{Example of tag suggestions when creating a post on the SuperUser site}
  \label{figSuperUserSuggestion}
\end{figure}

The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not public.
After exploratory testing, StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags.
The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.

So it stands to reason that the algorithm is relatively complex, and there could be possible scaling and efficiency concerns with the algorithm's implementation, so it has not yet been deployed to StackOverflow.
Alternatively, it is also possible that StackExchange decided to try out a tag recommendation system on a few of their smaller sites first, before deploying this type of system on their most popular StackOverflow site.
Nonetheless, it is encouraging that StackExchange finds tag recommendation models beneficial to the user and has started to add them to their question and answer sites.

\subsubsection{Google+}

The hashtag recommendation system deployed for the Google+ social media site is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.

\begin{figure}[!htbp]
  \scalebox{.4}{\resizebox{\linewidth}{!}{\includegraphics{google+HashtagExample-crop.pdf}}}
  \caption{Example post on Google+}
  \label{figGoogle+Hashtags}
\end{figure}

The workflow and user interface provided for recommending hashtags for Google+ is similar to how a hashtag recommendation system for Twitter might look.
When generating content on Google+, the system automatically associates the most likely relevant hashtags with that post.
For the example in Figure \ref{figGoogle+Hashtags}, the system recommended \emph{\#SFGiants} and \emph{\#BusterPosey} as hashtags for the user-created post.
The user is free to remove any or all of the tags that the system has associated with the post, and can also use tags that were not included in the recommended set.

Google+'s tag-recommendation algorithm is not public, but it appears to be rather sophisticated.
Looking at the example post in Figure \ref{figGoogle+Hashtags}, \emph{\#BusterPosey} is suggested without having that specific word in the post.
Also \textcite{GoogleKeynote2013} mentioned that the algorithm uses Google's sophisticated image recognition software to suggest hashtags for images that are included in the post (e.g., \emph{\#EiffelTower}),
even when the post does not have any keywords to provide additional context.

Google is apparently highly confident in the accuracy of their system.
On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
For Google+, the user interface automatically tags a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.

@@

A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
The latent topics for the collection of tweets were generated using a Latent Dirichlet Allocation (LDA) statistical methodology.
This is similar to a factor analysis approach in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance. 
Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
The LDA model performed well, and generated relevant topics.
When the model produced five most-likely hashtags for each tweet, 80\% of the time at least one of those hashtags was suitable (as determined by human raters).

However, a topic-generating model is not performing the same task as a hashtag-prediction model.
The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
It is unclear though if a user would want to label their tweet with these general topics.
Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
Now it is certainly the case that the topics generated by an LDA model may be similar to the hashtags used in the post.
However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
On the other hand, a topics-based approach is more general than a hashtags-based approach, since topics can be suggested for posts that do not use hashtags (which is quite common across Twitter).

@@

As some initial research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
To incorporate the user's prior tag history into the model, the base-level activation term ($B_{i}$) was modified from the original equation in ACT-R.
$B_{i}$ is based on log-odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
To incorporate a user's prior, a weighted average for $p_{i}$ was used,
where the probability of using tag $i$ ranges from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count increases.

The global probability is the same $p_{i}$ used in \textcite{Stanley2013}'s study.
The user's specific probability is computed the same way that the global probability is computed (i.e., the log odds that the tag occurred relative to the occurrence of all other tags).
However, each user's probability is constrained to the set of tags that the specific user has used in the past. 
When a user has no past tag history, the global prior probability is used.
Once a user has provided at least five tags, that user's specific prior probability is used.
Between zero and five tags, both terms are used but the weight of the global prior relative to the user prior decreases as the number of tags increases.

Both linear decay and exponential decay transition functions were tried.
Using either transition function, overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.

This research suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
in order to account for the fact that a user's tagging preferences change over time.

@@

\subsection{Handling Stop Words}

Stop words are usually removed by either filtering based on a predetermined list or filtering based on the high-frequency words that occur in the dataset.
\textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a predetermined list.

All three stop word techniques (i.e., predetermined list, high-frequency filtering, and entropy weighting) were explored for this analysis.

@@

\subsubsection{Visualizing Tag Use}

To better illustrate the dynamics of each user's tagging behavior,
profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figPriorHTByTime}.
The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
What is interesting is the large tag space for these users.
There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][StackOverflow]{\includegraphics[width=8cm]{HTByTime-520957-crop.pdf}}
    \hfill
    \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
    \caption{
      Tagging profile for a single user.
      The x-axis represents the specific hashtag a user chooses for a post and time progresses vertically.
      A point in the plot represents when a user chooses a specific hashtag for a specific post.
      A user's chosen hashtag for their oldest post that contains a hashtag is the lowest point on the plot.
    }
    \label{figPriorHTByTime}
  }%
\end{figure}

For example, the StackOverflow user in Figure \ref{figPriorHTByTime} has a pronounced band where the same hashtag is chosen repeatedly over time. 
However there are many instances of other tags being used in parallel to the tag that is producing the band (that is repeatedly being used). 
The Twitter user has a few clusters where the same hashtag is chosen repeatedly but only for a brief period of time.
As time progresses, a band of choosing another hashtag is observed, and then the user switches to a new hashtag after that.
More recent posts for the Twitter user are more varied, and there is no longer a single hashtag that the user chooses repeatedly.
This helps illustrate how challenging modeling this task can be.
The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is changing over time.

Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
These plots were chosen from the set to also show that the amount of tag variation for individual users also varies: some use many tags and some use just a few.
To accurately model human performance in this domain, the model must be able to make correct predictions across all variations of user tagging behavior.

\subsubsection{Visualizing Model Performance}

As a first method for understanding model performance, profile plots of model's predicted tags for individual users were overlaid on that user's actual tag use.
These plots for StackOverflow and Twitter are included in Figures \ref{figPriorHTMByTimeSO} and \ref{figPriorHTMByTimeT} respectively.
These are the same plots shown in Figure \ref{figPriorHTByTime} with model predictions overlaid in red.

\begin{figure}[!htbp]
  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
  \caption{Model tagging profile of a single user on StackOverflow}
  \label{figPriorHTMByTimeSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.7}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
  \caption{Model tagging profile of a single user on Twitter}
  \label{figPriorHTMByTimeT}
\end{figure}

Also shown in these figures is how model predictions change as the decay rate parameter changes.
With lower decay rate values, the model is less likely to switch from tag to tag for each post.
This makes sense because lower decay rate values lead to the model weighting frequency more than recency in the activation calculation.
These plots also suggest that model accuracy is best when a blend of frequency and recency is used (i.e., a decay rate between 0 and 20).
This will be more formally tested when model accuracy across decay rate values is analyzed across all users in each dataset slice and all dataset slices.

@@

Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{
    Model performance for a single dataset slice for StackOverflow.
    This includes model accuracy as a function of decay rate value for each user in the dataset slice.
    Model accuracy is shown both when the standard from of base-level activation ($B_{i}$) and the optimized learning form of the equation are used. 
  }
    \label{figPriorSOQSliceDsStd}
  }%
\end{figure}

@@

Chapter 5 will examine how model accuracy is influenced by two methodological concerns (adding an offset term and examining the effect of dataset subset for Twitter)
and five architectural concerns
(method of handling stop words, influence of dataset size, amount of compression used for the vector-based models, method to combine the prior and context for the vector-based models, and word order).
The two methodological concerns were examined to ensure that the technique to calibrate model parameters was accurate and also that results for Twitter were not specific to the subset of data collected.

The five architectural concerns were examined for the following reasons:
During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.

@@

\textcite{Sahlgren2008} showed how different levels of compression in the random-permutation model influence accuracy, so this was examined for three reasons:
to replicate the findings,
as an additional check to ensure that the random-permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.

@@

The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013}, so this was the first architectural concern that was examined.
\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
A data-driven frequency-weighted approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-generated stoplist hardly increased performance.
However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.
So alongside looking at frequency weighting, previously generated, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on entropy was also explored.

@@

\subsection{Amount of Compression for Random Permutation}

Another parameter in the random permutation model is the number of rows used to represent all of the different words.
Each word's environment vector is represented by two randomly placed ones and two negative ones across the rows.
All that is required is that the four positions and signs of each of these environment vectors is unique.
But this means multiple environment vectors can overlap when looking at a single row.
This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words.
As the number of rows decreases, the amount of cross talk between the environment vectors increases, compression increases, and the amount of information stored decreases.
As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.

Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
A separate co-occurrence representation was built for each of these row values.
Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.

Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure \ref{figContextDimSO}.
In the figure, the model name that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random permutation model for all other results.
The plot is also broken out for different size datasets used to build the co-occurrence matrix.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-dimBySizeSO-crop.pdf}}}
  \caption{
    Effect of compression on random permutation model for StackOverflow.
    Error bars represent the 95\% bootstrapped confidence interval for the mean.
  }
  \label{figContextDimSO}
\end{figure}

>>>>

@@

@@

The core question motivating this research is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
This process can be framed as a memory retrieval problem.
The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
That is, each user on a social media site has created content that contains a set of tags within the context of each post.

@@

@@

@@

@@

@@

@@

@@

@@

@@

Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.

@@

@@

Models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
Recommendation models for StackOverflow and Twitter will be described in more detail to follow.

@@

@@

@@

@@

@@

@@

Looking first at only a single dataset slice, model accuracy for all users in that slice was analyzed.
A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{
    Model performance for a single dataset slice for StackOverflow.
    This includes model accuracy as a function of decay rate value for each user in the dataset slice.
    Model accuracy is shown both when the standard from of base-level activation ($B_{i}$) and the optimized learning form of the equation are used. 
  }
    \label{figPriorSOQSliceDsStd}
  }%

\end{figure}

@@

Chapter 5 will examine how model accuracy is influenced by two methodological concerns (adding an offset term and examining the effect of dataset subset for Twitter)
and four architectural concerns
(method of handling stop words, influence of dataset size, method to combine prior and context, and word order).
The two methodological concerns were examined to ensure that the technique to calibrate model parameters was accurate and also that results for Twitter were not specific to the subset of data collected.

The four architectural concerns were examined for the following reasons:
During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.

@@

@@

The method of handling stop words has been shown to influence model accuracy \parencite{Sahlgren2008,Stanley2013,Jones2007}, so this was the first architectural concern that was examined.
\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random permutation model:
A data-driven frequency-weighted approach was tried, along with using a previously-generated stoplist, and no stop word removal at all.
Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-generated stoplist hardly increased performance.
However, \citeauthor{Sahlgren2008} did not try to weight the stop words and only tried methods to remove them.

All three stop word techniques (i.e., predetermined list, high-frequency filtering, and \textcite{Stanley2013}'s entropy weighting) were explored for this analysis.

@@
